Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-04-02 document/ms.tex
discussion of posterior samplings over catalogs

The ``likelihood function'' given in \eq{true-like} is a model of the
probability of observing a catalog with \emph{true} periods and radii.
If we want to include the effects of measurement uncertainties, we need to
take a moment to consider the meaning of measurement uncertainties.

\dfmtodo{Introduce the idea of catalogs with uncertainties being a description
of a posterior sampling over catalogs: $\catalog^{(n)} \sim
p(\catalog\,|\,\obs{\catalog},\,\interim)$.}

@@

\section{Separable distribution}

The first model that we consider for $\rate(\period,\,\radius)$ is a simple
separable distribution.
This assumption means that the rate can be written as
\begin{eqnarray}
\rate(\period,\,\radius) &=& \rate(\period)\,\rate(\radius)
\end{eqnarray}
and, more specifically, we model the period distribution as a broken power law
\dfmtodo{CHECK ME!!}
\begin{eqnarray}
\rate(\period) &\propto& \frac{\period^\alpha \,
(\period+\period_0)^{\beta-\alpha}}{\period_0^\beta}
\end{eqnarray}
and the radius distribution as a piecewise constant function (or histogram)
where the parameters are the bin heights.

\begin{figure}[htbp]
\begin{center}
\includegraphics{figures/separable-radius.pdf}
\end{center}
\caption{%
The inferred distribution of exoplanet radii assuming a separable rate
function.
The black lines are draws from the posterior probability distribution of bin
heights and the red points are previous measurements based on the same
dataset (\citealt{petigura}).
\figlabel{separable-radius}}
\end{figure}

\section{General distribution}

>>>>

The likelihood function given in \eq{true-like} is a model of the
probability of observing a catalog with \emph{true} periods and radii.
If we want to include the effects of measurement uncertainties, we need to
take a moment to consider the meaning of measurement uncertainties.

The catalog of planet candidates considered in this \paper\
(\citealt{petigura}) was generated by drawing samples from a posterior
probability distribution for physical parameters (period, radius, impact
parameter, \etc) conditioned on \kepler\ light curves.
The $k$th entry in this catalog is, therefore, a representation of the
posterior PDF
\begin{eqnarray}\eqlabel{measurement}
p(\entry_k\,|\,\obs{\entry}_k,\,\interim)
\end{eqnarray}
using the notation from the previous section.
In \eq{measurement}, the notations \interim\ indicates that the measurement
was made with a specific choice of prior over parameters
$p(\entry_k\,|\,\interim)$.
We will refer to these priors as \emph{interim priors} because our goal is to
infer the distribution that is consistent with the data.

More generally, a published catalog represents a \emph{posterior PDF over
plausible catalogs}
\begin{eqnarray}\eqlabel{catalog-pdf}
p(\catalog\,|\,\obs{\catalog},\,\interim) \quad.
\end{eqnarray}
It is normally safe to assume that each entry in the catalog is conditionally
independent given the choices of interim priors.
\dfmtodo{Why and when is this a good/bad/necessary assumption?}
Under this assumption, \eq{catalog-pdf} factorizes as
\begin{eqnarray}\eqlabel{catalog-factor}
p(\catalog\,|\,\obs{\catalog},\,\interim) &=&
\prod_{k=1}^K p(\entry_k\,|\,\obs{\entry}_k,\,\interim) \quad.
\end{eqnarray}

It will be useful in the following discussion to represent the catalog PDF as
a set of $N$ samples
\begin{eqnarray}\eqlabel{catalog-samples}
\catalog^{(n)} &\sim& p(\catalog\,|\,\obs{\catalog},\,\interim) \quad.
\end{eqnarray}
For the general catalog, these could be generated using MCMC or a similar
sampling algorithm.
Under the conditional independence assumption of \eq{catalog-factor}, however,
the problem is much simpler.
Each entry $\entry_k$ can be sampled independently and then the catalog
samples $\catalog^{(n)}$ are given by the outer product of the independent
sample lists.

@@

\section{Real data}
